---
title: "Exercise_1_KNN"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(gmodels)
library(class)
library(caret)
library(swirl)
```

## Exercise 2
#official saviour https://github.com/LennartOlsen/pca-digits/blob/master/main.r
## Excercise 2.4
This task is about reconstructing data using PCA. First using these functions we can plot an
image of a single cipher ( for plotting images do not convert id to data frame ):

PCA loading vector - the directions in feature space along which the data vary the most
PCA scores - projections along these directions


``` {r important functions}

# show the cipher gotten from reconstruction
show_pca_reconstruction <- function(dataset, searched_variance){
  
  limit_pca_rank <- get_pcs_for_accum_variance(dataset, searched_accum_var)# how to get the components that give us the % of variance?
  wanted_pcs <- prcomp(dataset[,(2:ncol(dataset))], center = TRUE,scale. = TRUE, rank. = limit_pca_rank)
  print(wanted_pcs)
  print(paste0("function returns data with dimensions", dim(wanted_pcs)))

  restr <- get_reconstructed_pca(wanted_pcs)
  
  #cipher_index <- 3*400 + 3 # get any index of a cipher - ideally show all 10 digits
  # find indexes of digits and print them
  #for (digit in c(0,9)){
  #  cipher_index <- 1
  #  while (dataset[cipher_index, ]$V1 != digit){
  #    cipher_index <- cipher_index + 1
  #  } 
  #  image(get_img_rot(restr, cipher_index), col = gray(0:100/100) )
  #}
  image(get_img_rot(restr, cipher_index), col = gray(0:100/100) )
}

# show the first 10 eigenvectors of PCA decomposition 
show_pca_eigenvectors <- function(dataset, searched_accum_var){
  limit_pca_rank <- get_pcs_for_accum_variance(dataset, searched_accum_var)# how to get the components that give us the % of variance?
  wanted_pcs <- prcomp(dataset[,(2:ncol(dataset))], center = TRUE,scale. = TRUE, rank. = limit_pca_rank)
  imageSize <- 18
  # Plot the first 10 eigenvectors and plot as images
  eigenVector <- 1:10
  #wanted_pcs$rotation[,eigenVector]
  for (e in eigenVector) {
    imageNewM <- matrix(wanted_pcs$rotation[,e],nrow = imageSize,ncol = imageSize,byrow = FALSE)
    print(image(imageNewM))
  }
}
```
``` {r }
load("../data/idList-cornered-100-2021.Rdata") 
id <- do.call(rbind, idList[1:6])
id <- as.data.frame(id)
id$V1 <- factor(id$V1)

shuffled_df <- id[sample(nrow(id)),]

cipher_index <- 100
image(get_img_rot(shuffled_df, cipher_index), col = gray(0:100/100) )
#accumulated_variance_v <- c(80, 90, 95, 99)
accumulated_variance_v <- c(80)
for (searched_accum_var in accumulated_variance_v){
  show_pca_reconstruction(shuffled_df, searched_accum_var)
  show_pca_eigenvectors(shuffled_df, searched_accum_var)
}
```
## 2.4.5. 
The last exercise is to compare the outcomes between two different ciphers. For instance, two
different ciphers, (e.g. row 43 and row 456 represent a ’0’ and a ’1’), compare the 10 first scores
and see if you can spot a difference. Try also to calculate the mean for all 400 instances of these
ciphers and compare the first 10 scores. Can you spot a pattern when comparing with the
loadings
``` {r }
library("stats")
# library(princomp)

load("../data/idList-cornered-100-2021.Rdata") 
id <- do.call(rbind, idList[1:6])
id <- as.data.frame(id)
id$V1 <- factor(id$V1)

shuffled_df <- id#[sample(nrow(id)),]

wanted_pcs <- prcomp(shuffled_df[,(2:ncol(shuffled_df))], center = TRUE,scale. = TRUE)
```
``` {r }

# compare the 10 first scores and see if you can spot a difference 
show_pca_eigenvectors_cipher <- function(dataset, cipher_number){
  # dataset <- dataset[cipher_number, ]
  wanted_pcs <- prcomp(dataset[,(2:ncol(dataset))], center = TRUE,scale. = TRUE)
  imageSize <- 18
  # Plot the first 10 eigenvectors and plot as images
  eigenVector <- 1:10
  for (e in eigenVector) {
    imageNewM <- matrix(wanted_pcs$rotation[,e],nrow = imageSize,ncol = imageSize,byrow = FALSE)
    print(image(imageNewM))
  }
}

show_ciphers <- c(43, 456)
for (cipher in show_ciphers){
  show_pca_eigenvectors_cipher(id, cipher)
}


```
OK tried to do function for the second part of the exercise
``` {r }

# DITCHED THE FUNCTION TO FIRST TRY ON PAPER
# # calculate the mean for all instances of these ciphers and compare the first 10 scores
# get_mean_rot_200_cipher <- function(){
#   # instead of "wanted_pcs$rotation[,e]" just put mean of pca$rotation[,e] of 400 ciphers
#   wanted_pcs <- prcomp(dataset[,(2:ncol(dataset))], center = TRUE,scale. = TRUE)
#   imageSize <- 18
#   
#   # digit 0 is from row 0 to row 200
#   zeros_rows <- c(1:200)
#   ones_rows <- c(201:400)
#   
#   rot_zero_v <- c()
#   wanted_pcs_zero <- prcomp(dataset[zeros_rows,(2:ncol(dataset))], center = TRUE,scale. = TRUE)
#   for (e in eigenVector) {
#     rot_zero_v <- c(rot_zero_v, wanted_pcs$rotation[,e])
#   }
#   
# }


# digit 0 is from row 0 to row 200
zeros_rows <- c(1:200)
ones_rows <- c(201:400)

rot_zero_v <- c()
wanted_pcs_zero <- prcomp(id[zeros_rows,(2:ncol(id))], center = TRUE,scale. = TRUE)
for (e in c(1:10)) {
  rot_now <-  wanted_pcs$rotation[,e]
  rot_zero_v <- c(rot_zero_v, mean(rot_now))
  
}
imageSize <- 18
imageNewM <- matrix(rot_zero_v,nrow = imageSize,ncol = imageSize,byrow = FALSE)
print(image(imageNewM))

```

``` {r their code}

# cipherNumber <- 43
# cumPic(10)
plot(wanted_pcs$x[43,1:10], ylim=c(-1,0.5), col="red")
# cipherNumber <- 456
# cumPic(10)
plot(wanted_pcs$x[456,1:10])
points(wanted_pcs$x[456,1:10], col="green")
biplot(wanted_pcs[1:10])
summary(wanted_pcs)

# cipher 2
cipherNumber <- 1001
wanted_pcs$x[cipherNumber, 1:10]
# cipher 0
cipherNumber <- 400
wanted_pcs$x[cipherNumber, 1:10]
#Mean
# cipher 2
no <- 1:10
result2 <- c()
result2_2 <- c()
for (a in no) {
  result2[a] <- mean(wanted_pcs$x[801:1200, a]) 
  result2_2[a] <- mean(wanted_pcs$rotation[, a])
}
result2
#result2_2

# cipher 0
result1 <- c()
for (a in no) {
  result1[a] <- mean(wanted_pcs$x[1:400, a])  
}
result1
```





