---
title: "Exercise_1_KNN"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(gmodels)
library(class)
library(caret)
library(swirl)
```

## Exercise 2

## Excercise 2.1 Principal Component Analysis (PCA)
Perform a PCA on the data for the “all persons in” and “disjunct” data set.
### 2.1.1 Show the standard deviation ( From prcomp Eigenvalues ), the proportion of variance and the cumulative sum of variance of the principal components. (In the report the first 10-20 principal components, should be sufficient to illustrate the tendencies.)


``` {r PCA}
print_plots_pca <- function(data){
  # help https://www.datacamp.com/community/tutorials/pca-analysis-r
  df_pca <- prcomp(data[,c(2:235)], center = TRUE,scale. = TRUE)
  pca_sum <- summary(df_pca)
  print(pca_sum)
  
  # Zouchi mentioned using squaring of sdev!!!
  
  pca_std_variance <- pca_sum$importance[1,1:20]
  plot(pca_std_variance, main="standard deviation of first 20 PCAs")
  
  pca_proportion_of_variance <- pca_sum$importance[2,1:20]
  plot(pca_proportion_of_variance, main="proportion of variance of first 20 PCAs")
  
  pca_cumulative_proportion <- pca_sum$importance[3,1:20]
  plot(pca_cumulative_proportion, main="cumulative proportion first 20 PCAs")

}


# All persons in
load("../data/idList-co-100.Rdata")
id <- do.call(rbind, idList[1:10])
id <- as.data.frame(id)
id$V1 <- factor(id$V1)
# id - huge dataframe with everyone as TRAINING

set.seed(423)
shuffled_df <- id[sample(nrow(id)),]
print_plots_pca(shuffled_df)

# disjunt data
id <- idList[[1]]
shuffled_df <- id[sample(nrow(id)),]
print_plots_pca(shuffled_df)

```

## 2.1.2 Show the performance of selecting enough principal components to represent 80%, 90%, 95%, 99% of the accumulated variance. For each test vary “k” in kNN, try 3 reasonable values.
``` {r performance_knn_with_pca}
# Zouchi told sth to use prcomp.x if I remember correctly - we really need to read the documentation
# how to get the data with only few pcas?
load("../data/id100.Rda")
shuffled_df <- id[sample(nrow(id)),]

df_pca <- prcomp(id[,c(2:235)], center = TRUE,scale. = TRUE)
pca_sum <- summary(df_pca)
print(pca_sum)

accumulated_variance_v <- c(80, 90, 95, 99)

```









Leftover from exc 1
```{r}
load("../data/id100.Rda")
set.seed(423)

k_error_avg_vec <- c()
timing_now_avg_vec <- c()
ks <- c( 19)

for (k_now in ks ){
  #print(paste0("k_now = ", k_now))
  k_error <- 0
  k_error_avg <- 0
  
  timing_now_avg <- 0
  for (i in 1:10) {
    # shuffle dataset
    shuffled_df <- id[sample(nrow(id)),]
    
    # split 10/90 for training and testing data
    test_df <- shuffled_df[1:200,]  # shuffled_df[1:360,]
    train_df <- shuffled_df[201:400,] # shuffled_df[361:400,]
    
    # get the labels - supervision part
    id_train_labels <- train_df[,1]
    id_test_labels <- test_df[,1]
    
    # check the time
    t_start <- Sys.time()
    # get the prediction
    numbers_test_pred <- knn(train = train_df[,c(2:325)], test = test_df[,c(2:325)], cl = id_train_labels, k=k_now)
    timing_now <- (Sys.time() - t_start ) * 1000 #ms
    
    k_error <- k_error +mean(id_test_labels != numbers_test_pred)
    timing_now_avg <- timing_now_avg + timing_now

  }
  # get summary of one k
  k_error_avg <- k_error / 10
  k_error_avg_vec <- c(k_error_avg_vec, k_error_avg)
  
  timing_now_avg <- timing_now_avg / 10
  timing_now_avg_vec <- c(timing_now_avg_vec, timing_now_avg) 
  
  print(paste0("k_now = ", k_now, " k_error_avg = ", k_error_avg, " timing_now_avg=", timing_now_avg ))
}
#plot(ks, k_error_avg_vec, type="o", ylab="misclassification error")
#plot(ks, timing_now_avg_vec, type="o", ylab="average execution time [ms]")

```

``` {r case A}
#DONT KNOW WHY IT GIVES ERRORS!

load("data/idList-co-100.Rdata")
id <- do.call(rbind, idList[1:10])
id <- as.data.frame(id)
id$V1 <- factor(id$V1)

# id - huge dataframe with everyone as TRAINING
train_df <- id
id_train_labels <- train_df[,1]

person_error_v <- c()
# individually test sets for each person
for (person_df in idList){
  print("new_person")
  test_df <- person_df
  id_test_labels <- person_df[,1]
  
  # get the prediction
  numbers_test_pred <- knn(train = train_df[,c(2:325)], test = test_df[,c(2:325)],cl = id_train_labels, k=15)
  
  person_error <- mean(id_test_labels != numbers_test_pred)
  person_error_v <- c(person_error_v, person_error)
} 

#print(paste0("dim(person_error_v) = ", dim(person_error_v)))
plot(1:length(idList), person_error_v, type="o", ylab="misclassification error", xlab="person number", main="case A")

```