---
title: "Exercise_1_KNN"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(gmodels)
library(class)
library(caret)
library(swirl)
```

## Exercise 1 – Data preparation and
KNN


```{r}
load("data/id100.Rda")


get_img <- function(df, sample_no){
  # get matrix of number from dataset
  # df-dataset; sample_no - which sample to return the matrix
  img <- data.matrix(df[sample_no,2:235])
  matrix(img, nrow = 18, ncol = 18, byrow = FALSE)
}

# visualize image
image(get_img(id,3), col = gray(0:100/100))
```

``` {r df variant knn}

load("data/id100.Rda")
set.seed(423)

# shuffle dataset
rows <- sample(nrow(id))
shuffled_df <- id[rows,]

# split 50/50 for training and testing data
test_df <- shuffled_df[1:200,]
train_df <- shuffled_df[201:400,]

id_train_labels <- train_df[,1]
id_test_labels <- train_df[,1]



numbers_test_pred <- knn(train = train_df, test = test_df,
cl = id_train_labels, k=21)

```
```{r}

CrossTable(x = id_test_labels, y = numbers_test_pred,
prop.chisq=FALSE)
```
```{r}
folds <- createFolds(id$X1, k = 10)

lapply(folds,function(x){
     holdout <- rep(FALSE,nrow(id))
     holdout[x] <- TRUE
     table(holdout,id$X1)
})
```

# TRY DIFFERENT K VALUES - calculate average from 10 tries of each K value, each time shuffling the data
``` {r try different ks}
 accuracy <- function(x){sum(diag(x)/(sum(rowSums(x)))) * 100}

load("data/id100.Rda")
set.seed(423)

k_accuracy_avg_vec <- c()
k_error_avg_vec <- c()
ks <- c(1, 5, 9, 15, 19, 25, 29)
# ks <- 1:40

for (k_now in ks ){
  #print(paste0("k_now = ", k_now))
  k_error <- 0
  k_error_avg <- 0
  
  k_accuracy <- 0
  k_accuracy_avg <- 0
  
  for (i in 1:10) {
    # shuffle dataset
    shuffled_df <- id[sample(nrow(id)),]
    
    # split 10/90 for training and testing data
    test_df <- shuffled_df[1:200,]  # shuffled_df[1:360,]
    train_df <- shuffled_df[201:400,] # shuffled_df[361:400,]
    
    # get the labels - supervision part
    id_train_labels <- train_df[,1]
    id_test_labels <- test_df[,1]
    
    # get the prediction
    numbers_test_pred <- knn(train = train_df, test = test_df,
    cl = id_train_labels, k=k_now)

    # create table to calculate accuracy
    tab <- table(numbers_test_pred,id_test_labels)
    accuracy_now <- accuracy(tab)
    #print(paste0("i=",i ,"  accuracy_now=", accuracy_now))
    
    k_error <- mean(id_test_labels != numbers_test_pred)
    k_accuracy <- k_accuracy + accuracy_now
  }
  # get summary of one k
  k_accuracy_avg <- k_accuracy / 10
  k_accuracy_avg_vec <- c(k_accuracy_avg_vec, k_accuracy_avg)
  
  k_error_avg <- k_error / 10
  k_error_avg_vec <- c(k_error_avg_vec, k_error_avg)
  
  print(paste0("k_now = ", k_now, "  k_accuracy_avg = ", k_accuracy_avg, " k_error_avg = ", k_error_avg ))
}
plot(ks, k_accuracy_avg_vec, type="o", ylab="accuracy")
plot(ks, k_error_avg_vec, type="o", ylab="misclassification error")
```
# Using cross-validation to find best k
``` {r cross validation}
# help https://genomicsclass.github.io/book/pages/crossvalidation.html - seems like for every fold check all the ks
library(class)
library(caret)
load("data/id100.Rda")

# shuffle dataset
shuffled_df <- id[sample(nrow(id)),]

# for every fold check all the ks
#ks <- c(1,5,9,13,17,21,25,29)
#ks <- c(1, 35, 71)
#ks <- 1:40
ks <- c(15)

mse_k_v <- c()  # mse per one k
for (k in ks){
  # do the folds
  folds <- createFolds(shuffled_df$X1, k = 10)
  
  mse_kf <- 0 # mean squared error for all the folds for this k 
  
  for (fold in folds){
    #current fold - to ignore in training data but to use as a validation dataset: fold -> test!
    df_fold <- shuffled_df[ fold , ]
    
    # use not-fold data as training and fold data as test
    not_f_df <- shuffled_df[ -fold , ]
    f_df <- shuffled_df[ fold,  ]
    
    # get labels 
    f_train_labels <- not_f_df[,1]
    f_test_labels <- f_df[,1]
    
    # predict with knn
    f_test_pred <- knn(train = not_f_df[,c(2:325)], test = f_df[,c(2:325)], cl = f_train_labels, k=k)
    
    # calculate error
    f_error <- mean(f_test_labels != f_test_pred)
    print(paste0("      k =", k, " f_error =", f_error))
    mse_kf <- mse_kf + f_error
  }
  print(paste0("k = ", k, " mse_kf = ", mse_kf))
  mse_k_v <- c(mse_k_v, mse_kf/10)
}
print(paste0("mse_k_v = ", mse_k_v))
plot(ks, mse_k_v, type="o", ylab="missclassification error")

```
It seems like the loops should be inverted, so thats why we want to do this:

Error? Why the plots look like they look? Is there a mistake with shuffling or sth like that? It says that no matter what k we are using, the error rates are extremely low.
``` {r inverted loops}
# help https://genomicsclass.github.io/book/pages/crossvalidation.html - seems like for every fold check all the ks
library(class)
library(caret)
load("data/id100.Rda")

# shuffle dataset
set.seed(423)
shuffled_df <- id[sample(nrow(id)),]

# for every fold check all the ks
ks <- c(1,5,9,13,17,21,25,29)
mse_k_all_folds_avg <- rep(0, length(ks)) # vector for storing error of each k 

# do the folds
folds <- createFolds(shuffled_df$X1, k = 10)

fold_no <- 0 # just for printing the progress

#current fold - to ignore in training data but to use as a validation dataset: fold -> test!
for (fold in folds){
  fold_no <- fold_no + 1
  print(paste0("fold_no =", fold_no))
  
  fold <- sample(fold) # shuffle id numbers of fold 
  
  # use not-fold data as training and fold data as test
  not_f_df <- shuffled_df[ -fold, ]
  f_df <- shuffled_df[ fold, ]
  
  # get labels 
  f_train_labels <- not_f_df[,1]
  f_test_labels <- f_df[,1]
  
  # predict with knn - test all k's for each fold

  k_no <- 0 # just to keep track
  k_error <- 0
  for (k in ks){
    k_no <- k_no + 1
    f_test_pred <- knn(train = not_f_df[,c(2:325)], test = f_df[,c(2:325)], cl = f_train_labels, k=k)
    
    k_error <- mean(f_test_labels != f_test_pred) # error of current k in this fold
    print(paste0("      k =", k, " k_error =", k_error))
    print( mse_k_all_folds_avg) # Error vector that we want to plot
    
    # we just want to plot the results so we came up with idea of saving the errors in vector
    # for each k in the vector all the folds are summed up together, and then we get the average
    mse_k_all_folds_avg[k_no] <- mse_k_all_folds_avg[k_no] + k_error
  }
}
mse_k_all_folds_avg <- mse_k_all_folds_avg / length(folds)

print(paste0("mse_k_all_folds_avg = ", mse_k_all_folds_avg))
plot(ks, mse_k_all_folds_avg, type="o", ylab="missclassification error", main="each k had checked the folds")
```