---
title: "Exercise_1_KNN"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(gmodels)
library(class)
library(caret)
library(swirl)
```

## Exercise 1 – Data preparation and
KNN
```{r test}
apple <- c('red', 'green', "yellow")
print(class(apple))

```


```{r}
load("data/id100.Rda")


get_img <- function(df, sample_no){
  # get matrix of number from dataset
  # df-dataset; sample_no - which sample to return the matrix
  img <- data.matrix(df[sample_no,2:235])
  matrix(img, nrow = 18, ncol = 18, byrow = FALSE)
}

# visualize image
image(get_img(id,3), col = gray(0:100/100))
```
## get another df full of images?
Should we have dataset full of imgs? - for sure then easy to shuffle
we will have to compute the distances - easier when having an image

``` {r}
# get list of matrices with images
numbers_list <- list()
for ( i in 1:400){
  num_mat <- get_img(id,i)
  numbers_list[[(length(numbers_list) + 1)]] <- num_mat # append list
}

```

```{r list variant}
# shuffle list
shuffled_list <- sample(numbers_list)
```

Verify shuffling of the list
```{r}
image(numbers_list[[1]], col = gray(0:100/100))

image(shuffled_list[[1]], col = gray(0:100/100))
```
Split to training set and testing set 50/50
```{r}


```


END HERE
``` {r df variant}

load("data/id100.Rda")
set.seed(423)

# shuffle dataset
rows <- sample(nrow(id))
shuffled_df <- id[rows,]

# split 50/50 for training and testing data
test_df <- shuffled_df[1:200,]
train_df <- shuffled_df[201:400,]

id_train_labels <- train_df[,1]
id_test_labels <- train_df[,1]

numbers_test_pred <- knn(train = train_df, test = test_df,
cl = id_train_labels, k=21)

```
```{r}

CrossTable(x = id_test_labels, y = numbers_test_pred,
prop.chisq=FALSE)
```
```{r}
folds <- createFolds(id$X1, k = 10)

lapply(folds,function(x){
     holdout <- rep(FALSE,nrow(id))
     holdout[x] <- TRUE
     table(holdout,id$X1)
})
```

# WORKING PLOT HERE
# TRY DIFFERENT K VALUES - calculate average from 10 tries of each K value, each time shuffling the data
``` {r try different ks}
 accuracy <- function(x){sum(diag(x)/(sum(rowSums(x)))) * 100}

load("data/id100.Rda")
set.seed(423)

k_accuracy_avg_vec <- c()
k_error_avg_vec <- c()
ks <- c(1, 5, 9, 15, 19, 25, 29)
# ks <- 1:40

for (k_now in ks ){
  #print(paste0("k_now = ", k_now))
  k_error <- 0
  k_error_avg <- 0
  
  k_accuracy <- 0
  k_accuracy_avg <- 0
  
  for (i in 1:10) {
    # shuffle dataset
    shuffled_df <- id[sample(nrow(id)),]
    
    # split 10/90 for training and testing data
    test_df <- shuffled_df[1:200,]  # shuffled_df[1:360,]
    train_df <- shuffled_df[201:400,] # shuffled_df[361:400,]
    
    # get the labels - supervision part
    id_train_labels <- train_df[,1]
    id_test_labels <- test_df[,1]
    
    # get the prediction
    numbers_test_pred <- knn(train = train_df, test = test_df,
    cl = id_train_labels, k=k_now)

    # create table to calculate accuracy
    tab <- table(numbers_test_pred,id_test_labels)
    accuracy_now <- accuracy(tab)
    #print(paste0("i=",i ,"  accuracy_now=", accuracy_now))
    
    k_error <- mean(id_test_labels != numbers_test_pred)
    k_accuracy <- k_accuracy + accuracy_now
  }
  # get summary of one k
  k_accuracy_avg <- k_accuracy / 10
  k_accuracy_avg_vec <- c(k_accuracy_avg_vec, k_accuracy_avg)
  
  k_error_avg <- k_error / 10
  k_error_avg_vec <- c(k_error_avg_vec, k_error_avg)
  
  print(paste0("k_now = ", k_now, "  k_accuracy_avg = ", k_accuracy_avg, " k_error_avg = ", k_error_avg ))
}
plot(ks, k_accuracy_avg_vec, type="o", ylab="accuracy")
plot(ks, k_error_avg_vec, type="o", ylab="misclassification error")
```
``` {r trying intro from https://genomicsclass.github.io/book/pages/crossvalidation.html }

# shuffle dataset
shuffled_df <- id[sample(nrow(id)),]

# split 10/90 for training and testing data
test_df <- shuffled_df[1:200,]  # shuffled_df[1:360,]
train_df <- shuffled_df[201:400,] # shuffled_df[361:400,]

# get the labels - supervision part
id_train_labels <- train_df[,1]
id_test_labels <- test_df[,1]

# get the prediction
numbers_test_pred <- knn(train = train_df, test = test_df,cl = id_train_labels, k=9)

error <- mean(id_train_labels != numbers_test_pred)
print(paste0("error = ", error))


# now try to do the folds
folds <- createFolds(id$X1, k = 2)  # idx
#folds <- createFolds(id, k = 10)  # idx online they use not the labels but full dataset
print(sapply(folds, length))  # each fold has 400 samples


#rbind()

for (fold in folds){
  print("fold")
  df_fold <- shuffled_df[ fold , ]
  print(dim(df_fold))
  print(dim(shuffled_df))
  
  # test_df is always from the fold
  # train_df is the all the other folds
  
  df_not_fold <- shuffled_df - df_fold
  print(dim(df_not_fold))

  
  f_train_df <-df_fold[1:360,]  # 90% training
  f_test_df <-df_fold[361:400,]  # 10% test
  
  f_train_labels <- f_train_df[,1]
  f_test_labels <- f_test_df[,1]
  
  # but shouldnt we use data that doesnt contain the current fold???
  f_test_pred <- knn(train = f_train_df, test = f_test_df,
    cl = f_train_labels, k=100)
  
  f_error <- mean(f_test_labels != f_test_pred)
  print(paste0("f_error = ", f_error))
  
  
}
#testing <- id[ folds[[1]] , ]

```

``` {r Ines}
library(class)
library(caret)

# shuffle dataset
shuffled_df <- id[sample(nrow(id)),]

# split 10/90 for training and testing data
test_df <- shuffled_df[1:200,]  # shuffled_df[1:360,]
train_df <- shuffled_df[201:400,] # shuffled_df[361:400,]

# get the labels - supervision part
id_train_labels <- train_df[,1]
id_test_labels <- test_df[,1]

# get the prediction
numbers_test_pred <- knn(train = train_df, test = test_df,cl = id_train_labels, k=9)

error <- mean(id_train_labels != numbers_test_pred)
print(paste0("error = ", error))


# now try to do the folds
folds <- createFolds(id$X1, k = 2)  # idx
#folds <- createFolds(id, k = 10)  # idx online they use not the labels but full dataset
print(sapply(folds, length))  # each fold has 400 samples

for (fold in folds){
  print("fold")
  df_fold <- shuffled_df[ fold , ]
  #print(df_fold)
  
  f_train_df <-df_fold[1:360,]  # 90% training
  f_test_df <-df_fold[361:400,]  # 10% test
  
#current fold = fold to ignore in training data but to use as validation dataset
#other folds = training data 
  
  f_train_labels <- f_train_df[,1]
  f_test_labels <- f_test_df[,1]
  not_f_df_labels <- not_f_df[, 1]
  
  # but shouldnt we use data that doesnt contain the current fold???
  not_f_df <- f_train_df[ -fold , ]
  f_df <- f_test_df[ fold,  ]
  
  f_test_pred <- knn(train = not_f_df, test = f_df, cl = not_f_df[, 1], k=21)
  
  f_error <- mean(f_test_labels != f_test_pred)
  print(paste0("f_error = ", f_error))
  
}

```



``` {r}
# y - labels
y <- id_train_labels
X <- train_df
# knn(train =  X, test = X, cl=y, k=5)

idx <- createFolds(y, k=10)
sapply(idx, length)

the_labels <- knn(train =  X, test = X, cl=y, k=5)

#1st fold of data
#shuffled_df[folds[1]]
```

# USE K-FOLDS to find best K
``` {r https://genomicsclass.github.io/book/pages/crossvalidation.html}
load("data/id100.Rda")
set.seed(423)

ks <- 1:12  # set k from 1 to 12
res <- sapply(ks, function(k) { # apply a function to a whole dataset
  ##try out each version of k from 1 to 12
  res.k <- sapply(seq_along(idx), function(i) {
    ##loop over each of the 10 cross-validation folds
    ##predict the held-out samples using k nearest neighbors
    pred <- knn(train=Xsmall[ -idx[[i]], ],
                test=Xsmall[ idx[[i]], ],
                cl=y[ -idx[[i]] ], k = k)
    ##the ratio of misclassified samples
    mean(y[ idx[[i]] ] != pred)
  })
  ##average over the 10 folds
  mean(res.k)
})


```


``` {r Erick's code}
library(class)
nor <-function(x) { (x -min(x))/(max(x)-min(x))   }
#load("id100.Rda")
id_norm <- as.data.frame(lapply(id[,c(2:325)], nor))
ran <- sample(1:nrow(id), 0.5 * nrow(id)) 
id_train <- id_norm[ran,]
id_test <- id_norm[-ran,]
id_target_category <- id[ran,1]
id_test_category <- id[-ran,1]
library(class)
pr <- knn(id_train,id_test,cl=id_target_category,k=1)
tab <- table(pr,id_test_category)
accuracy <- function(x){sum(diag(x)/(sum(rowSums(x)))) * 100}
accuracy(tab)
nor <-function(x) { (x -min(x))/(max(x)-min(x))   }
#THIS SHOULD WORK : 
#load("id100.Rda")
ran <- sample(1:nrow(id), 0.5 * nrow(id)) 

id_train <- id_norm[ran,]
id_test <- id_norm[-ran,]
id_target_category <- id[ran,1]
id_test_category <- id[-ran,1]

pr <- knn(id_train,id_test,cl=id_target_category,k=1)
tab <- table(pr,id_test_category)
accuracy <- function(x){sum(diag(x)/(sum(rowSums(x)))) * 100}
accuracy(tab)
```
Higher K lower precision


## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
